---
output: github_document
bibliography: creating-maps.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Welcome to cs7

This is the online, up-to-date home for the course *Intermediate and Advanced R for Spatial Data Analysis*.

It is also a *metapackage* in the sense that when you install it, many of the packages the tutorials depend on will be installed on your computer. Because this metapackage is not on CRAN, we use the **devtools** package to install it:

```{r, eval=FALSE}
if(!require(devtools))
  install.packages("devtools")
devtools::install_github("robinlovelace/cs7")
```

## Learning outcomes

By the end of the course participants will:

- Be more efficient general R programmers
- Understand standard (non-spatial) data structures and how to subset, process and analyse them
- Understand the structure of spatial data in R
- Be able to query, subset and analyse spatial objects
- Have a working knowledge of fundamental GIS functions such as changing projections
- Be proficient in the use of R to create maps using add-on packages such as tmap
- Have some experience with advanced functionality such as raster data and interactive maps

## Course content

### Day 1: Intermediate R 

- (Re)introduction to RStudio (9:30 - 11:15)
      - 1 hr lecturing, 45 minutes practical
      - Practical handout: [creating-maps-in-R](https://github.com/Robinlovelace/Creating-maps-in-R), Part I

- Objects, functions and concepts for efficient R programming (11.30 - 13.45) 
      - 30 minutes intro, 45 minutes hands on, 20 minutes lecture, 40 minutes practical
      - Practical handout: [Objects, functions and concepts for efficient R
programming](https://www.dropbox.com/s/w9ol2by7b7civr6/Objects_Functions.pdf?dl=0), by Lex Comber
      - Exercise: Read sections [4.1](https://csgillespie.github.io/efficientR/efficient-set-up.html#top-5-tips-for-an-efficient-r-set-up) to [4.2](https://csgillespie.github.io/efficientR/efficient-set-up.html#operating-system) of @colin_gillespie_robin_lovelace_efficient_2016 and complete the Exercise at the end of 4.2.

**13.45 - 14.30 Lunch**

- Data manipulation (14:30 - 16:00) 
    - Practical: work through [Chapter 5](https://csgillespie.github.io/efficientR/efficient-data-carpentry.html) of *Efficient R Programming*

- Plotting paradigms  (16:15 - 17:45) 
      - Lecture and practical handout: [creating-maps-in-R](https://github.com/Robinlovelace/Creating-maps-in-R) (Part IV)
    - **Base graphics**
    - **ggplot2**
    - **tmap**
    - **leaflet**
    - **mapview**

## Day 2: Basics of R as a GIS

- Spatial data in R, sp classes and projections (09.00 -11.15) 
      - Practical handout 1: [Spatial data in R](https://www.dropbox.com/s/9eozr5slpj638es/Manipluating_Spatial_Objects.pdf?dl=0) by Lex Comber
      - Practical handout 2: [creating-maps-in-R](https://github.com/Robinlovelace/Creating-maps-in-R), Part II)
    
- Loading, plotting and interrogating spatial data, including shapefiles, xy and spatial queries (11:30 - 13:45) 
    - (practical handout: [creating-maps-in-R](https://github.com/Robinlovelace/Creating-maps-in-R), Part III)

**13.45 - 14.30 Lunch**

- Manipulating spatial objects with a focus on rgeos (14:30 - 16:00)
    - Course handout: [Manipulating Spatial Objects](https://www.dropbox.com/s/9eozr5slpj638es/Manipluating_Spatial_Objects.pdf?dl=0) by Lex Comber

- Introduction raster data with R (16:15 - 17:45) 
      - Course handout: [Introduction to Raster Analysis](https://www.dropbox.com/home/Teaching/rmaterials-lc?preview=Raster_Analysis.pdf) by Lex Comber

## Day 3: Advanced R for spatial data analysis

- Raster/vector operations with R - class conversions and aggregation with raster (09:00 - 11:15)

<!-- - Spatio-temporal data with **spacetime** (11:30 - 13:45) -->

- Transport data with **stplanr** (11:30 - 13:45)
    - Exercise: Work through the package's vignette (`vignette("introducing-stplanr")`)

**13.45 - 14.30 Lunch**

- Points pattern analysis - points to surfaces by IDW, Kriging (14:30 - 16:00) 
    - Handout: [Point Patter analysis](https://www.dropbox.com/s/gn1i21rsvipbwom/Point_Pattern.pdf?dl=0) by Lex Comber

- Advanced graphics: online mapping with **mapview** and leaflet and static maps with **tmap** (16:15 - 17:45)

## Prerequisites and course home

Working knowledge of R and RStudio is assumed.

Background reading (available online):

- @torfs_very_2014
- @colin_gillespie_robin_lovelace_efficient_2016
- @lovelace_introduction_2014

Background reading (not available online):

- @bivand_applied_2013
- @dorman_learning_2014
- @lamigueiro_displaying_2014
- @brunsdon_introduction_2015

## References

```{r, echo=FALSE, eval=FALSE}
library(RefManageR)
bib = ReadZotero(user = 216746, .params = list(collection = 'VZ45CJWD', key = Sys.getenv("ZOTERO"), limit = 20))
WriteBib(bib = bib, file = "creating-maps.bib")
```

